---
title: "Class10-29"
author: "Tabea"
date: "29 Oktober 2018"
output: html_document
---
---
title: "In class assignment week 2, part 2. A worked example using control flow (for loops, if statements, etc)"
author: "Ian Dworkin"
output: 
  html_document:
    keep_md: yes
    number_sections: yes
---

# Introduction
Let's do a little exercise integrating some of the things we have learned. Here are some Illumina HiSeq reads for one of our recent projects:

```{r}
read_1 <- "CGCGCAGTAGGGCACATGCCAGGTGTCCGCCACTTGGTGGGCACACAGCCGATGACGAACGGGCTCCTTGACTATAATCTGACCCGTTTGCGTTTGGGTGACCAGGGAGAACTGGTGCTCCTGC"

read_2 <- "AAAAAGCCAACCGAGAAATCCGCCAAGCCTGGCGACAAGAAGCCAGAGCAGAAGAAGACTGCTGCGGCTCCCGCTGCCGGCAAGAAGGAGGCTGCTCCCTCGGCTGCCAAGCCAGCTGCCGCTG"

read_3  <- "CAGCACGGACTGGGGCTTCTTGCCGGCGAGGACCTTCTTCTTGGCATCCTTGCTCTTGGCCTTGGCGGCCGCGGTCGTCTTTACGGCCGCGGGCTTCTTGGCAGCAGCACCGGCGGTCGCTGGC"
``


#Question 1. what species are these sequences from?
They are from Drosophila melanogaster.

#Question 2. Put all three of these reads into a single object (a vector).  What class will the vector `reads` be? Check to make sure! How many characters are in each read (and why does `length()` not give you what you want.. try...)
reads <- c(read_1, read_2, read_3)
The class of the vector is "character". 
nchar() gives 124 characters for each read. length() only shows the number of elements, not the number of characters.


#Question 3. Say we wanted to print each character (not the full string) from read_1, how do we do this using a for loop? You may wish to look at a function like `strsplit()` to accomplish this (there are other ways.)

Replace the blanks.
```{r}
read_1_split <- strsplit(___, split = ___, fixed = T) 
```
for (x in read_1) {print(strsplit(x, split = character(0), fixed = T))}

#Question 4. What kind of object does this return? How might we make it a character vector again?

A list is returned.
read_1_char <- as.character(unlist(read_1_split))

#Question 5. How about if we wanted the number of occurrences of each base? Or better yet, their frequencies? You could write a loop, but I suggest looking at the help for the `table()` function... Also keep in mind that for most objects `length()` tells you how many elements there are in a vector. For lists use `lengths()` (so you can either do this on a character vector or a list, your choice)

table(read_1_char)
read_1_char
 A  C  G  T 
10 41 43 30 

site_freq <- table(read_1_split)/lengths(read_1_split)
print(site_freq, digits = 2)
#Länge des Vektors wir mit einbezogen -> hier Länge = 1, wichtig bei längeren Vektoren


#Question 6. How would you make this into a nice looking function that can work on either  a list or vectors of characters? (Still just for a single read)
if (mode(x) == "list") {tab <- table(x)/lengths(x)}

#Question 7. Now how can you modify your approach to do it for an arbitrary numbers of reads? You could use a loop or use one of the apply like functions (which one)?
if (lentgh(x) == 1 & mode(x) == "character"){
  x <- strsplit(x, split = "", fixed = T)
  x <- as.character(unlist(x))
  }
if (mode(x) == "list") {
  tab <- table(x)/lengths(x)}
  else { 
  tab <- table(x)/length(x)
  }
  return(tab)
  }
#Question 8. Can you revise your function so that it can handle the input of either a string as a single multicharacter vector, **or** a vector of individual characters **or** a list? Try it out with the vector of three sequence reads (`reads`).  
basefreq <- sapply(reads, BaseFrequencies, USE.NAMES = F)
print(basefreq, digits = "")